---
title: "HW 11"
author: "Mohammad Mottahedi"
date: "April 11, 2016"
output: 
    pdf_document:
      highlight: zenburn
      toc: yes
---

# Problem 1

\subsection{a}

```{r}
temp <- read.ftable(textConnection("Z z1 z2
Y y1 y2 y3
X 
x1 125 7 11 5 106 18
x2 124 6 22 3 109 9
x3 146 6 0 2 111 0"))

df <- data.frame(temp)


# saturated model
model <- glm(Freq ~ X + Y + Z + X*Y + X*Z + Y*Z + X*Y*Z, data = df, family = poisson(link = "log"))
summary(model)


# homogenous model
model <- glm(Freq ~ X + Y + Z + X*Y + X*Z + Y*Z, data = df, family = poisson(link = "log"))
summary(model)


# independece model
model <- glm(Freq ~ X + Y + Z, data = df, family = poisson(link = "log"))
summary(model)

# joint indep model (XY,Z)
model <- glm(Freq ~ X + Y + Z + X*Y, data = df, family = poisson(link = "log"))
summary(model)

# joint independece model (XZ,Y)
model <- glm(Freq ~ X + Y + Z + X*Z, data = df, family = poisson(link = "log"))
summary(model)

# joint independece model (YZ,X)
model <- glm(Freq ~ X + Y + Z + Y*Z, data = df, family = poisson(link = "log"))
summary(model)

# cond independece model (XY,XZ)
model <- glm(Freq ~ X + Y + Z + X*Z + X*Z, data = df, family = poisson(link = "log"))
summary(model)

# cond independece model (XY,XZ)
model <- glm(Freq ~ X + Y + Z + X*Z + X*Z, data = df, family = poisson(link = "log"))
summary(model)

# cond independece model (XY,YZ)
model <- glm(Freq ~ X + Y + Z + X*Y + Y*Z, data = df, family = poisson(link = "log"))
summary(model)

# cond independece model (XZ,YZ)
model <- glm(Freq ~ X + Y + Z + X*Z + Y*Z, data = df, family = poisson(link = "log"))
summary(model)
```
The best model is conditional independece model (XY,YZ).

The estimated errors are high.

dropping the cells with zero counts.
```{r}
df = df[df$Freq != 0,]

# saturated model
model <- glm(Freq ~ X + Y + Z + X*Y + X*Z + Y*Z + X*Y*Z, data = df, family = poisson(link = "log"))
summary(model)


# homogenous model
model <- glm(Freq ~ X + Y + Z + X*Y + X*Z + Y*Z, data = df, family = poisson(link = "log"))
summary(model)


# independece model
model <- glm(Freq ~ X + Y + Z, data = df, family = poisson(link = "log"))
summary(model)

# joint indep model (XY,Z)
model <- glm(Freq ~ X + Y + Z + X*Y, data = df, family = poisson(link = "log"))
summary(model)

# joint independece model (XZ,Y)
model <- glm(Freq ~ X + Y + Z + X*Z, data = df, family = poisson(link = "log"))
summary(model)

# joint independece model (YZ,X)
model <- glm(Freq ~ X + Y + Z + Y*Z, data = df, family = poisson(link = "log"))
summary(model)

# cond independece model (XY,XZ)
model <- glm(Freq ~ X + Y + Z + X*Z + X*Z, data = df, family = poisson(link = "log"))
summary(model)

# cond independece model (XY,XZ)
model <- glm(Freq ~ X + Y + Z + X*Z + X*Z, data = df, family = poisson(link = "log"))
summary(model)

# cond independece model (XY,YZ)
model <- glm(Freq ~ X + Y + Z + X*Y + Y*Z, data = df, family = poisson(link = "log"))
summary(model)

# cond independece model (XZ,YZ)
model <- glm(Freq ~ X + Y + Z + X*Z + Y*Z, data = df, family = poisson(link = "log"))
summary(model)
```


\subsection{b}


```{r}
delta1 <- c(0,0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  0, 1, 0, 0, 0)
delta2 <- c(0,0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,  0, 0, 0, 0, 1)

df <- data.frame(temp)


# saturated model
model <- glm(Freq ~ X + Y + Z + X*Y + X*Z + Y*Z + X*Y*Z + delta1 + delta2, data = df, family = poisson(link = "log"))
summary(model)

# homogenous model
model <- glm(Freq ~ X + Y + Z + X*Y + X*Z + Y*Z + delta1 + delta2, data = df, family = poisson(link = "log"))
summary(model)


# independece model
model <- glm(Freq ~ X + Y + Z, data = df, family = poisson(link = "log"))
summary(model)

# joint indep model (XY,Z)
model <- glm(Freq ~ X + Y + Z + X*Y + delta1 + delta2, data = df, family = poisson(link = "log"))
summary(model)

# joint independece model (XZ,Y)
model <- glm(Freq ~ X + Y + Z + X*Z+ delta1 + delta2, data = df, family = poisson(link = "log"))
summary(model)

# joint independece model (YZ,X)
model <- glm(Freq ~ X + Y + Z + Y*Z+ delta1 + delta2, data = df, family = poisson(link = "log"))
summary(model)

# cond independece model (XY,XZ)
model <- glm(Freq ~ X + Y + Z + X*Y + X*Z+ delta1 + delta2, data = df, family = poisson(link = "log"))
summary(model)

# cond independece model (XY,YZ)
model <- glm(Freq ~ X + Y + Z + X*Y + Y*Z+ delta1 + delta2, data = df, family = poisson(link = "log"))
summary(model)

# cond independece model (XZ,YZ)
model <- glm(Freq ~ X + Y + Z + X*Z + Y*Z+ delta1 + delta2, data = df, family = poisson(link = "log"))
summary(model)

```


# Problem 2

```{r}
temp <- read.ftable(textConnection("TonsilSize Normal Slightl-enlarged Very-enlarged
CarrierCond
Carrier 19 29 24
Non-Carrier 497 560 269"))

df <- data.frame(temp)

options(contrast=c("contr.treatment","contr.poly"))

#independece model
model <- glm(Freq ~ CarrierCond + TonsilSize, data = df, family = poisson(link = "log"))

summary(model)


xCarrier <- rep(c(1,2), c(3,3))
ySize <- rep(c(1,2,3), 2)

#linear by linear 
model <- glm(Freq ~ CarrierCond + TonsilSize + xCarrier * ySize,
             data = df, family = poisson(link = "log"))

summary(model)
```

the linear by linear model fits better compared to independet model.

both, hw4 and linear by linear model show that being a carrier and tonsil is related.

# Problem 3

\subsection{a}



```{r}

count <- c(425, 17, 80, 36, 10, 555, 74, 47, 7, 34, 771, 33, 5, 14, 29, 452)

at04 <- rep(c("N", "MW", "S", "W"), c(4,4,4,4))
at16 <- rep(c("N", "MW", "S", "W"), 4)

iN <- (at04 == "N") * (at16 == "N")
iMW <- (at04 == "MW") * (at16 == "MW")
iS <- (at04 == "S") * (at16 == "S")
iW <- (at04 == "W") * (at16 == "W")


at04_N <- (at04 == "N")
at04_MW <- (at04 == "MW")
at04_S <- (at04 == "S")
at04_W <- (at04 == "W")


at16_N <-  (at16 == "N")
at16_MW <- (at16 == "MW")
at16_S <-  (at16 == "S")
at16_W <-  (at16 == "W")

symm1 <- 1 * (at04 == "N") * (at16 == "N")
symm2 <- 2 * (at04 == "MW") * (at16 == "MW")
symm3 <- 3 * (at04 == "S") * (at16 == "S")
symm4 <- 4 * (at04 == "S") * (at16 == "S")

symm5 <- 5 * (at04 == "N") *  (at16 == "MW") + 5 * (at04 == "MW") * (at16 == "N")
symm6 <- 6 * (at04 == "S")  * (at16 == "N") + 6*   (at04 == "N") *(at16 == "S")
symm7 <- 7 * (at04 == "W") * (at16 == "N") + 7 *(at04 == "N") * (at16 == "W")
symm8 <- 8 * (at04 == "S")*  (at16 == "MW")+ 8 * (at04 == "MW")* (at16 == "S")
symm9  <- 9 *(at04 == "MW") * (at16 == "W") + 9  * (at04 == "W")* (at16 == "MW")
symm10  <- 10 *(at04 == "S") * (at16 == "W")+ 10 * (at04 == "W")* (at16 == "S")

symm=symm3+symm1+symm4+symm6+symm2+symm5+symm6 + symm7 + symm8 + symm9 + symm10

df <- data.frame(at04, at16, count, iN, iMW, iW , iS , symm)

df
```

### symmetry model


```{r}
model=glm(count~symm1+symm4+symm6+symm2+symm5+symm3+symm7 + symm8+ symm9+ symm10,data=df,family=poisson(link=log))
summary(model)


    


```


$df= 6$ 
, $G^2 = 134.45$ , p-value $= `r 1- pchisq(134.45,6)`$

symmetry model doesnot seem to be a good fit.


###  Quasi-symmetry Model

```{r}

model=glm(count~at04_N + at04_W+ at04_MW+ at04_S + at16_N+ at16_W+ at16_MW+ at16_S +symm1+symm4+symm6+symm2+symm5+symm3+symm7 + symm8+ symm9+ symm10,data=df,family=poisson(link=log))
summary(model)
```
$df=3$ 
, $G^2 =  3.9324$ , p-value $= `r 1- pchisq( 3.9324,3)`$

the quasi-symmetry model fits moderatlely well.

\subsection{b}

$G^2(marginal homogenety) = G^2(symmtery) - G^2(quasi-symmtery) = `r 134.45 - 3.932`$

$df=3$ 

p-value= `r 1-pchisq(134.45 - 3.932, 3)`$

the marginal homgeneity model does not fit the data.



