---
title: "hw1"
author: "Mohammad Mottahedi"
date: "January 13, 2016"
output: pdf_document
---

```{r}
set.seed(1234)
```

__Problem 1__ (16 points). Which scale of measurement is most appropriate
for the following variables?

(a) Treatment group (treatment or control) in a clinical trial?

answer: Binary variable.

(b) Age range in a clinical trial?

answer: Continuous variables grouped into small number of categories.

(c) Geographic location (North America, Southeast Asia, Europe)?

answer: Nominal.

(d) Michelin stars?

answer: Ordinal.


__Problem 2__ (16 points). Suppose in a certain area, the number of ships
per 10 by 10 kilometer region of the ocean has a Poisson distribution with
parameter 0.2.

(a) What is the probability of observing 5 ships in one 10 by 10 kilometer
region?

$p(y) = \frac{e^{-\mu}\mu^y}{y!}=\frac{e^{-0.2} 0.2^ 5}{5!}= `r dpois(5, 0.2)`, \mu = 0.2, y=5$


(b) What is the probability of observing no more than 5 ships in a 100 by
100 kilometer region?

There are 100 (independent) 10x10 $km^2$ region in a 100x100 $km^2$  region, number of ships in each of the 100 regions has possion distrubiton $\mu = 0.2$ $(X_i)$. The sum $X_1 +...+ X_{100}$ has the poisson distribuiton with mean $\mu_{total} = \mu_1 + ... + \mu_{100} =  `r 100 * 0.2`$.

$p(y)= p(y=0) + p(y=2) + p(y=3) + p(y=4) + p(y=5) =  `r sum(dpois(0:5, 20))`$


__Problem 3__ (18 points). When the 2000 General Social Survey asked subjects
whether they would be willing to accept cuts in their standard of living to
protect the environment, 344 of 1170 subjects said Yes. Note: Some helpful
R and SAS code is provided on ANGEL. See hw1. R and hw1_ help. SAS .

(a) Estimate the population proportion who would say Yes.

estimated population proportion: $\hat{p}= \frac{344}{1170} = `r 344/1170`$

estimated standard error: $\hat{SE} = \frac{\sqrt{\hat{p} (1 - \hat{p})}}{\sqrt{n}} = `r sqrt( ((344/1170) * (1 - (344/1170))) / 1170)`$

(b) Conduct significance test to determine whether a majority of the population
would say Yes. Report and interpret the p-value.

$H_0: p <= 0.5$

$H_A: p> 0.5$

$Z = \frac{\overline{p} - 0.5}{\hat{SE}}= \frac{0.2940171 - 0.5}{0.0133196}= `r (0.2940171 - 0.5)/0.0133196`$

${p-value} = `r 1 - pnorm((0.2940171 - 0.5)/ 0.0133196)` > 0.05$

it's very likely to see value as low as 0.29 so we can't reject the null hypothesis. 

(c) Construct and interpret 99% confidence interval for the population proportion
who would say Yes. Here you should construct a classical confi-
dence interval, i.e. based on Wald test; later we will see some problems
with this approach

$\hat{p}= \frac{344}{1170} = `r 344/1170`$

$\hat{SE} = \frac{\sqrt{\hat{p} (1 - \hat{p})}}{\sqrt{n}} = `r (sqrt( (344/1170) * (1 - (344/1170)))) /sqrt(1170)`$

(`r 0.2940171 - qnorm(.99) * 0.0133196` , `r 0.2940171 + qnorm(.99) * 0.0133196`)



__Problem 4__ (20 points). Each of 100 multiple-choice questions on an exam
has four possible answers, one of which is correct. For each question, a
student guesses by selecting an answer randomly.

(a) Specify the distribution of the students number of correct answers.

is binomial with probability mass fucntion
$p(y) = \frac{100!}{[y!(100-y)!]} 0.25^y (0.75)^{100-y}$

(b) Find the mean and standard deviation of that distribution. Would it be
surprising if the student made at least 50 correct responses? Why?

$E(Y)= n \pi = 100(\frac{1}{4}) = 25$ , $\sigma = \sqrt{n \pi(1 - \pi)} = \sqrt{ 100 \frac{1}{4}(1 - \frac{1}{4})} = `r sqrt(100 * 1/4 * 3/4)`$ 

Hypothesis testing:

$H_0: \mu \ge 50$,
$H_A: \mu < 50$

$Z = \frac{25 - 50}{4.330127/\sqrt(100)} = `r (25 - 50)/ (4.330127 / sqrt(100))`$

$p-value = `r pnorm((25 - 50)/ (4.330127 / sqrt(100)) )` < 0.05$ , yes it's suprising and it's unlikely to see 25 correct answers if the null hypothesis is correct, so we reject the null hypothesis.

(c) Specify the distribution of $(n_1 , n_2 , n_3 , n_4 )$ where $n_j$ is the number of
times the student picked choice j.

Multinomial distrubiton 

$p(n_1, n_2, n_3, n_4) = (\frac{n!}{n_1! n_2! n_3! n_4!}) \pi_1^{n_1} \pi_2^{n_2} \pi_3^{n_3} \pi_4^{n_4}$


(d) Find the mean of n j , the variance of n j , the covariance between n j and
n k and the correlation between n j and n k .

$E(n_j) = n \pi_j = 100(1/4) = 25$

$var(n_j) = n \pi_j (1 - \pi_j) = `r 100 * 1/4 * 3/4`$

$$\mathbf{cov} = \left[\begin{array}
{rrrr}
18.75 & -6.25 & -6.25 & -6.25 \\
-6.25 & 18.75 & -6.25 & -6.25 \\
-6.25 & -6.25 & 18.75 & -6.25 \\
-6.25 & -6.25 & -6.25 & 18.75
\end{array}\right]
$$



__Problem 5__ (30 points). Let X be the number of Chargers (San Diegos
NFL team) fans observed in a random sample of n = 30 graduate students
at Penn State. It is reasonable to assume that X ~ Bin(30; p). The true
proportion p is unknown, but it is likely to be small (especially in comparison
to potential number of Pittsburgh Steelers fans).
Note: Some helpful R and SAS code is provided on ANGEL; see hw1. R
and hw1_ help. sas . For plotting likelihood there is only R code.

(a) Assume for now that p = 0.04. Find the mean, the variance, and the
standard deviation of X. Find the probability for each of the following
events X = 0, X = 1, X = 2, X = 3, and X >= 4. You can do this by
hand, or use R or SAS.

$E(X) = np = `r 0.04 * 30`$,
$var(X) = np(1-p) = `r (30 * 0.04 * (1 - 0.04))`$

$p(X=0) = `r dbinom(0,30,0.04)`$

$p(X=1) = `r dbinom(1,30,0.04)`$

$p(X=2) = `r dbinom(2,30,0.04)`$

$p(X=3) = `r dbinom(3,30,0.04)`$

$p(X \ge 4) = `r sum(dbinom(4:30,30,0.04))`$

(b) The classic (Wald) approximate 95% confidence interval for p given in
the most textbooks is 

$\hat{p} \pm 1.96 \sqrt{\frac{\hat{p}(1-\hat{p})}{n}}$

where $\hat{p} = X/n$. When does this interval become degenerate (i.e. when
are the lower and the upper bounds are the same)? If the true p was
actually 0.04, could this interval actually cover the true parameter 95%
of the time? Why or why not? (Hint: based on part (a), how often
would the interval become degenerate at zero?)

The interval is degenerate when $\hat{p}$ is close to 0 or 1. $\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}=0 \to \hat{p}(1-\hat{p})=0$.


with $p=0.04$ the 95% confidence interval is $(`r 0.04 - sqrt(0.04 *(1- 0.04) / 30)` , `r 0.04 - sqrt(0.04 *(1- 0.04) / 30)`)$ and does not cover the 0.04.




(c) Suppose that we observe two Chargers fans in the sample. Plot the
log-likelihood function for p over a range of values from p = 0.01 to
p = 0.20. Find the ML estimate and the value of the loglikelihood at $\hat{p}$.
Calculate the approximate 95% confidence interval for p based on the
expected information.

$l(p|x) = xlog(p) + (n-x)log(1-p)$

$X=2, n=30$

$l(p|x) = 2log(p) + (28)log(1-p)$

```{r}
p = seq(from=0.01, to=0.2, by=0.01)
y = 2 * log(p) + 28 * log(1-p)
plot(p, y, xlab="", ylab="log-likelihood", type="l")
```

log-likelihood at $l(\hat{p} = `r 2/30`) = `r 2 * log(2/30) + 28 * log(1 - 2/30)`$ 


$I(p) = \frac{n}{p(1-p)}$

95% confidence interval for p: $\hat{p} \pm 1.96 \sqrt{\frac{\hat{p}(1-\hat{p})}{n}} = (`r 2/30 - 1.96 * sqrt(2/30 * (1 - 2/30) / 30)` , `r 2/30 + 1.96 * sqrt(2/30 * (1 - 2/30) / 30)` )$



(d) Plot the log-likelihood function with respect to the transformed parameter over a range of  values from. Comment on this plot in comparison to the one in part (b).

(e) What is the 95% confidence interval for based on the expected infor-
mation, and then transform the end points back to scale of p.

(f) Now find an approximate 95% confidence interval based on the LR
method, and compare it to the two other intervals calculated above.
Which one is the most appropriate and why?

