---
title: "Hw8"
author: "Mohammad Mottahedi"
date: "March 20, 2016"
output: pdf_document
---
```{r}
set.seed(1234)
setwd('~/Dropbox/spring2016/discretDataAnalysis/hw_solutions/hw8/')
library(VGAM)
library(MASS)

```


# Problem 1

\subsection{part a}
    
```{r}
gator = read.table("gator.txt",header=T)
gator$Size = factor(gator$Size,levels=levels(gator$Size)[2:1])
totaln=sum(gator[1:16,5:9]) ## total sample size
rown=c(1:16)
for (i in 1:16) {
rown[i]=sum(gator[i,5:9]) 
rown
}

##sets Hancock as the baseline level
contrasts(gator$Lake)=contr.treatment(levels(gator$Lake),base=2)
#contrasts(gator$Lake)

##sets "small" as the refernce level
contrasts(gator$Size)=contr.treatment(levels(gator$Size),base=2)
#contrasts(gator$Size)


##sets male as the reference level
contrasts(gator$Gender)=contr.treatment(levels(gator$Gender),base=2)
#contrasts(gator$Gender)

fit = vglm(cbind(Bird,Invertebrate,Reptile,Other,Fish)~Lake+Size, data=gator, family=multinomial)
summary(fit)

```
The overal fit of the model can be evaluiated using residual deviance $52.4$ with $44$ degrees of freedom (p-value: $`r 1-pchisq(52.4,44)`$). Overal the fit of the model is not good.

model equations:

$log \frac{\pi_{food,lsg}}{\pi_{F,lsg}} = \alpha_{food} + \beta_{food,l}^{lake} + \beta_{food,s}^{size}$


$log \frac{\pi_{bird,lsg}}{\pi_{F,lsg}} = -2.03 -1.35 Oklawaha + 0.39 Trafford - 0.695 George + 0.6307 Large$

$log \frac{\pi_{Invertebrate,lsg}}{\pi_{F,lsg}} = -1.7492 +2.5956 Oklawaha + 2.7803 Trafford - 0.695 George -1.4582 Large$

$log \frac{\pi_{Invertebrate,lsg}}{\pi_{F,lsg}} = -2.4230 +1.2161 Oklawaha + 1.6925 Trafford - 1.2428 George +0.3513 Large$

$log \frac{\pi_{Invertebrate,lsg}}{\pi_{F,lsg}} = -0.7465 -0.8205 Oklawaha + 0.6902 Trafford - 0.8262 George -0.3316 Large$
    
    
\subsection{part b}
    
```{r}
denom = 1 + exp(-2.03 - 1.35  + 0.6307 ) + 
    exp(-1.7492 + 2.5956 - 1.4582 ) + 
    exp(-2.4230 + 1.2161 + 0.3513 ) + 
    exp(-0.7465 - 0.8205 - 0.3316 )
```
$P(fish|Large,...) = `r 1/denom`$

```{r}
denom = 1 + exp(-2.03 - 1.35) + 
    exp(-1.7492 + 2.5956) + 
    exp(-2.4230 + 1.2161) + 
    exp(-0.7465 - 0.8205)

```
$P(fish|small,...) = `r 1/denom`$

probability of fish being the main food source is 1.7 times (46%) for large aligators compared to small aligators (0.26)

\subsection{part c}
    
```{r}
gator = read.table("alligator.dat",col.names = c('lake', 'sex', 'size', 'food','count'))
#gator$size = factor(gator$size,levels=c("small", "large"))
gator$food = factor(gator$food, levels = c('fish', 'invertebrate', 'reptile', 'bird', 'other'))
result = polr( food ~ size + lake, weights=count, data=gator)
null_model = polr( food ~ 1, weights=count, data=gator)
sat_model = polr(food ~ size + lake + sex + size:lake + size:sex + sex:lake, weights=count, data=gator )

summary(result)

# comparing with null model 
anova(result, null_model)

# comparing with sat model
anova(result, sat_model)

 pchisq(deviance(result), df.residual(result), lower.tail = F)
```    
The model is significant compared to both null and saturated model which shows that this model at leat has one non-zero coefficeint. The p-value for the current model is close to zero so the model is a good fit.

The effect of size is oposite of what we found in part b. the probability of eating fish is greater for smaller aligators.


# Problem 2
    
__part a__

We have to fit $r-1$ equation which is equal to $3$.
The $3$ parameter coefficients are the same and the 3 interceprt are different.

__part b__

$x_1:$ decrease

$x_2:$ increase

$x_3:$ increase

__part c__

$x_1:$ much less

$x_2:$ not at all true

$x_3:$ strongly disagree


# Problem 3


$logit[P(Y \le j|x_1)] - logit[P(Y \le j|x_2)]$

$= log \frac{P(Y \le j|x_1) / P(Y > j|x_1)}{P(Y \le j|x_2) / P(Y > j|x_2)}$

$log[(1- \alpha - \beta x_2) (\alpha + \beta x_1) - (1 - \alpha - \beta x_1)(\alpha +\beta x_2)]$

$=\beta(x_1-x_2)$

if $x_1 - x_2=1$

$=\beta$


\subsection{part b}

from part a:

$log \frac{P(Y \le j|x_1) / P(Y > j|x_1)}{P(Y \le j|x_2) / P(Y > j|x_2)} = \beta(x_1-x2)$

the odds of being below $\le j$  at $x_1$ is $e^{\beta (x_1-x_2)}$ times the odds at $x_2$. 

for x=1,2 and j=1,2

|    | j1  |  j2 | 
|----- |-----|-----|
|x1  | p(y <= 1 / x1) | p(y >= 2 / x1)|
|x2  | p(y <= 1 / x2) | p(y >= 2 / x2)|
